My title

### 2024-09-20 paper/paper.qmd
delete civilian casualties

# Read data, https://stackoverflow.com/questions/78894145/quarto-document-not-seeing-project-directory-for-targets-data-functions-or-scr
raw_data <- read_csv(here("data/raw/toronto_fire_incidents.csv"))
cleaned_data <- read_csv(here("data/analysis/cleaned_fire_data_simple.csv"))
cleaned_data_tfs <-
  cleaned_data |>
  select(
    tfs_alarm_time, 
    tfs_arrival_time, 
    time_diff,
    civilian_casualties, 
    estimated_dollar_loss,
  )

@@

The second group of data shows the money loss, cause of fire and the presence of fire prevention equipments. @tbl-clean-cat shows the first five rows of the second group of data. "Estimated Loss in Dollar" is the estimated loss measured in dollar. "Area of origin" indicates the area fire originate. "Ignition Source" shows the object causing fire. "Fire Alarm Status", "Smoke Alarm Status" and "Sprinkler System Status" indicates the presence and operation of fire protection equipments.

head(cleaned_data_cat, 5) |>
  kable(
    col.names = c("Estimated Loss in Dollar", "Area of Origin", "Ignition Source", "Fire Alarm Status"
    , "Smoke Alarm Status", "Sprinkler System Status"),
      )

cat("Note: 'PO' = System present, \n'N' = System not present, \n'P' = System present but not operated")
```

@@

The first group of data shows the responsiveness of Toronto Fire Services. An example of this dataset is presented in @tbl-clean-tfs. "Alarm Time" is the time when TFS are notified of the incident. "TFS Arrival time" is the timestamp of first arriving unit. The difference in minute is calculated at the data cleaning step as "TFS Response Time". The number under "Civilian Casualties" indicates the fire related injury or fatality. "Estimated Loss in Dollar" is the estimated loss measured in dollar.

@@

head(cleaned_data_tfs, 5) |>
  kable(
    col.names = c("Alarm Time", "TFS Arrival Time", "TFS Response Time"
    , "Civilian Casualties", "Estimated Loss in Dollar"),
      )
```

@fig-clean-tfs-diff-dollar-time-1 shows the line plot of Toronto Fire Service's response time over the period between 2013 to 2023.

@@

cleaned_data_tfs |>
  ggplot(mapping = aes(x = tfs_alarm_time, y = estimated_dollar_loss)) + 
  geom_line() +
  theme_minimal() +
  labs(
    x = "Alarm Time", y = "Estimated dollar loss",
  )

@@

>>>>

# Read data, https://stackoverflow.com/questions/78894145/quarto-document-not-seeing-project-directory-for-targets-data-functions-or-scr
raw_data <- read_csv(here("data/raw/toronto_fire_incidents.csv"))
cleaned_data <- read_csv(here("data/analysis/cleaned_fire_data_simple.csv"))
cleaned_data_tfs <-
  cleaned_data |>
  select(
    tfs_alarm_time, 
    tfs_arrival_time, 
    time_diff,
    estimated_dollar_loss,
  )

@@

@@

The first group of data shows the responsiveness of Toronto Fire Services. An example of this dataset is presented in @tbl-clean-tfs. "Alarm Time" is the time when TFS are notified of the incident. "TFS Arrival time" is the timestamp of first arriving unit. The difference in minute is calculated at the data cleaning step as "TFS Response Time". "Estimated Loss in Dollar" is the estimated loss measured in dollar.

@@

head(cleaned_data_tfs, 5) |>
  kable(
    col.names = c("Alarm Time", "TFS Arrival Time", "TFS Response Time", "Estimated Loss in Dollar"),
      )
```

@fig-clean-tfs-diff-dollar-time-1 shows the line plot of Toronto Fire Service's response time over the period between 2013 to 2023. The response time fluctuated around a relatively consistent level, typically between 5 and 10 minutes, with occasional spikes where the response time increased significantly above the usual range, particularly around 2014 and 2019. @fig-clean-tfs-diff-dollar-time-2 shows the line plot of estimated dollar loss of the fire incidents. Most of the incidents resulted in relatively low dollar losses, but there are several spikes that represent incidents where the loss was significantly higher. Some of the most notable spikes occurred between 2015 and 2019, with losses exceeding 2 million dollars.

@@

cleaned_data_tfs |>
  ggplot(mapping = aes(x = tfs_alarm_time, y = estimated_dollar_loss)) + 
  geom_line() +
  theme_minimal() +
  labs(
    x = "Alarm Time", y = "Estimated dollar loss",
  )
```

@tbl-clean-tfs-summary-1 indicates the summary statistic and @fig-clean-tfs-boxplot-1 graphs the boxplot of TFS's response time. The value mainly range from 4.03 minutes to 5.7 minutes as shown in @tbl-clean-tfs-summary-1, with a mean value of 4.911868 minutes. Notably, in @fig-clean-tfs-boxplot-1, several outliers are indicated by dots above the upper whisker, suggesting occasional instances of significantly higher response times.

@tbl-clean-tfs-summary-2 indicates the summary statistic and @fig-clean-tfs-boxplot-2 depicts the distribution of financial losses in boxplot. In @tbl-clean-tfs-summary-2, values span from 0 dollar to over 3000000 dollars, with median value of 3000 dollars, mean value of 24551.64 dollars and standard deviation of 93520.08 dollars, indicating a skewed distribution. @fig-clean-tfs-boxplot-2 reveals numerous outliers above the upper whisker, highlighting instances of exceptionally high financial losses.

@@

The second group of data shows the money loss, cause of fire and the presence of fire prevention equipments. @tbl-clean-cat shows the first five rows of the second group of data. "Estimated Loss in Dollar" is the estimated loss measured in dollar. "Area of origin" indicates the area fire originate. "Ignition Source" shows the object causing fire. "Fire Alarm Status", "Smoke Alarm Status" and "Sprinkler System Status" indicates the presence and operation of fire protection equipments. In the second group of data, 'PO' = System present, 'N' = System not present, and 'P' = System present but not operated.

head(cleaned_data_cat, 5) |>
  kable(
    col.names = c("Estimated Loss in Dollar", "Area of Origin", "Ignition Source", "Fire Alarm Status"
    , "Smoke Alarm Status", "Sprinkler System Status"),
      )
```
